Clustering feature points

### 2013-04-19 text/clustering.tex
Added the feature detector directory

Both methods suffer from two shortcomings. Firstly we have that any 
defined group in one image might not correspond to a group in another 
image.  This means that we might end up in a situation where a good 
match between two images is scored low because most of it's geometric 
neighbors are in an adjecent group. Secondly if we intent to match the 
different groups as in \cite{wu2009bundling} we might end up matching 
two mismatching groups. To improve geometric matching I propose pooling 
the keypoints from two images into one set and partitioning them into 
groups together.  The advantages of this approach is that groups are 
consistent across images, and that keypoints from one image that don't 
have equivalents in the other image will simply be grouped in their own 
group which can easily be discarded during the matching. This is 
important for scene recognition where people might obscure details in 
the picture. In addition the proposed method generalizes to three or 
more images.  This makes it possible to easily assemble a set of 
partially overlapping pictures without any required knowledge of their 
position.  What's more, if one of the pictures in there is not part of 
the set, the features of the odd image out will not form clusters with 
the features of the other images and automatically be sorted out.

@@

\begin{algorithm}
\caption{Louvain-Match Algorithm}
\label{alg-simple}
\begin{algorithmic}
\Require $images$ : set of images, $n$ : int $\in \mathbb{N}$
\State $matches\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features of all images
	\State $f_i\gets getFeatures(I_i)$
\EndFor
\State $A\gets getAdjecencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets 1 - normalize(A)$
\Comment Shorter distance is higher weight
\State $A_{pruned}\gets pruneEdges(A_{norm},\alpha)$
\State $P\gets cluster(A_{pruned})$ 
\Comment P is a set of partitions
\ForAll{$p \in P$} \Comment p is a set of feature points
	\State $trait\gets \left\{p \cap f_i \mid p \cap f_i \neq 
\varnothing, f_i \in \left\{f_1, f_2,\; \ldots \;, f_n\right\}\right\}$
	\If{$\left\vert trait \right\vert > 1$}
		\State $matches\gets getPositions(trait)$
		\Comment Position of trait in each image
	\EndIf
\EndFor \\
\Return matches
\end{algorithmic}
\end{algorithm}

Algorithm \ref{alg-simple} shows the basic steps taken to find matches 
in two or more images. This algorithm is a simplified version of the 
proposed method completely agnostic of the geometry of the feature 
points in the picture. To better explain the algorithm here is an 
overview over the inputs and functions mentioned and how they work:
\begin{itemize}
\item[]{\emph{images}: The images we are matching}
\item[]{\emph{$n$}: The number of weights per row in the adjecency 
matrix}
\item[]{\emph{getFeatures}: Returns a set of feature points given an 
image. The type of feature point is not theoretically important, but in 
practice the BRIEF\cite{calonder2010brief} feature point detector and 
descriptor has worked well} 
% TODO: include a comparison of SURF, SIFT, etc
\item[]{\emph{getAdjecencyMatrix}: Based on the distance measure used to 
compare the similarity of the feature points a symmetric adjecency 
matrix is returned}
\item[]{\emph{pruneEdges}: The result of the Louvain clustering depends 
on the density of the graph. To ensure an optimal number of partitions 
we only keep the $n$ highest weights per row and column in the adjecency 
matrix, making sure it's still symmetric afterwards. In practice setting 
$n$ to around four or five usually yields good results}
\item[]{\emph{cluster}: Cluster the graph corresponding to the adjecency 
matrix given using the Louvain algorithm in \cite{blondel2008}}
\item[]{\emph{getPositions}: Given a trait consisting of features in 
several different images, find the positions of traits in each image and 
return them. This could either be done by taking a mean of the feature 
points in an image, or by returning the position of the highest matching 
feature points within the trait. In practice ... TODO}
\end{itemize}

\begin{algorithm}
\caption{Louvain-Match Algorithm With Geometry}
\label{alg-geometry}
\begin{algorithmic}
\Require $images$ : set of images, $\alpha$ : float $\in \left[0, 
1\right]$, $\beta$ : float $\in \left[0, 1\right]$
\State $matches\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features of all images
	\State $f_i\gets getFeatures(I_i)$
\EndFor
\State $A\gets getAdjecencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $B\gets getGeomDistanceMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $M\gets getInterImageMask(f_1, f_2,\; \ldots \;, f_n)$
\State $B_{norm}\gets 1 - normalize(G)$ \State $A_{norm}\gets 1 - 
normalize(A)$
\State $A_{geom}\gets A_{norm}\left[ \neg M\right] + \beta \cdot 
B_{norm}\left[M\right]$
\State $A_{pruned}\gets pruneEdges(A_{geom},\alpha)$
\State $P\gets cluster(A_{pruned})$ 
\ForAll{$p \in P$}
	\State $trait\gets \left\{p \cap f_i \mid p \cap f_i \neq 
\varnothing, f_i \in \left\{f_1, f_2,\; \ldots \;, f_n\right\}\right\}$
	\If{$\left\vert trait \right\vert > 1$}
		\State $matches\gets getPositions(trait)$
		\Comment Position of trait in each image
	\EndIf
\EndFor \\
\Return matches
\end{algorithmic}
\end{algorithm}

While clustered feature points are often geometrically close together, 
there might be cases where several points in the same image are similar 
enough to all be included in one cluster despite their geometrical 
distance.  To solve this problem we can either modify the clustering 
algorithm to take the geometric distances into account like 
\cite{das2008event}.  Alternatively we can modify the weight matrix to 
penalize putting distanced feature points in the same cluster. However 
we only know the geometrical distance between two feature points as long 
as both belong to the same image. One way to approach this problem is to 
replace all the weights of edges between two feature points within the 
same image with their distance. For the case of two images the original 
adjecency matrix is defined as follows:
%
$$A=
\begin{bmatrix}
w_{1,1}   & w_{1,2}   & \ldots & w_{1,n-1}   & w_{1,n}     \\
w_{2,1}   & w_{2,2}   & \ldots & w_{2,n-1}   & w_{2,n}     \\
\vdots    & \vdots    & \ddots & \vdots      & \vdots      \\
w_{n-1,1} & w_{n-1,2} & \ldots & w_{n-1,n-1} & w_{n-1,n}   \\
w_{n,1}   & w_{n,2}   & \ldots & w_{n,n-1}   & w_{n,n}
\end{bmatrix}=
\begin{bmatrix}
A_{im1} & S       \\
T       & A_{im2} \\
\end{bmatrix}
$$
%
In a similar manner we can define an adjecency matrix based on the 
geometrical distance. Note how for example $d_{1,n}$ doesn't contain any 
sensible information.
%
$$B=
\begin{bmatrix}
d_{1,1}   & d_{1,2}   & \ldots & d_{1,n-1}   & d_{1,n}     \\
d_{2,1}   & d_{2,2}   & \ldots & d_{2,n-1}   & d_{2,n}     \\
\vdots    & \vdots    & \ddots & \vdots      & \vdots      \\
d_{n-1,1} & d_{n-1,2} & \ldots & d_{n-1,n-1} & d_{n-1,n}   \\
d_{n,1}   & d_{n,2}   & \ldots & d_{n,n-1}   & d_{n,n}
\end{bmatrix}=
\begin{bmatrix}
B_{im1} & R       \\
Q       & B_{im2} \\
\end{bmatrix}
$$
%
Given these two and a weight factor $\beta$ we can construct a new 
weight matrix as follows:
%
$$C=
\begin{bmatrix}
\beta \cdot B_{im1} & S       \\
T       & \beta \cdot B_{im2} \\
\end{bmatrix}
$$
%
Since we prune the matrix afterwards, it's important to choose a weight 
factor that doesn't lead to all edges being either intra image edges or 
intra image edges, since the first would create no clusters across 
images and the latter would mean that the geometrical information would 
be discarded and the bipartite graph constituting of edges between 
feature points in the two images would be used instead.

>>>>

Both methods suffer from two shortcomings. Firstly we have that any 
defined group in one image might not correspond to a group in another 
image.  This means that we might end up in a situation where a good 
match between two images is scored low because most of it's geometric 
neighbors are in an adjecent group. Secondly if we intent to match the 
different groups as in \cite{wu2009bundling} we might end up matching 
two mismatching groups. To improve geometric matching I propose pooling 
the keypoints from two images into one set and partitioning them into 
groups together.  The advantages of this approach is that groups are 
consistent across images, and that keypoints from one image that don't 
have equivalents in the other image will simply be grouped in their own 
group which can easily be discarded during the matching. This is 
important for scene recognition where people might obscure details in 
the picture. In addition the proposed method generalizes to three or 
more images.  This makes it possible to easily assemble a set of 
partially overlapping pictures without any required knowledge of their 
position.  What's more, if one of the pictures is not part of the set, 
the features of the odd image out will not form clusters with the 
features of the other images and automatically be sorted out.

\section{Literature review}

Since the introduction of the SIFT\cite features{lowe2004sift}, 
collecting feature points from an image and using the descriptors to 
match images have been used in a plethora of different applications. The 
development of feature points have not been standing still since Lowe 
introduced the SIFT feature back in 2004. Various modifications and 
improments both to how features are collected and how features are best 
described have been suggested since then. In locating feature points in 
an image the focus

To cluster similar parts in the images in to partitions we need to 
describe these points in a way which is invariant (or close to 
invariant) to changes in light, image quality, size, as well as 
geometric transformations due to change in the camera position.

TODO!!!! A lot more needs to be written here ...

@@

\begin{algorithm}
\caption{Cluster-Match Algorithm}
\label{alg-simple}
\begin{algorithmic}
\Require $images$ : set of images, $n$ : int $\in \mathbb{N}$
\State $matches\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features of all images
	\State $f_i\gets getFeatures(I_i)$
\EndFor
\State $A\gets getAdjecencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $A_{norm}\gets 1 - normalize(A)$
\Comment Shorter distance is higher weight
\State $A_{pruned}\gets pruneEdges(A_{norm},\alpha)$
\State $P\gets cluster(A_{pruned})$ 
\Comment P is a set of partitions
\ForAll{$p \in P$} \Comment p is a set of feature points
	\State $trait\gets \left\{p \cap f_i \mid p \cap f_i \neq 
\varnothing, f_i \in \left\{f_1, f_2,\; \ldots \;, f_n\right\}\right\}$
	\If{$\left\vert trait \right\vert > 1$}
		\State $matches\gets getPositions(trait)$
		\Comment Position of trait in each image
	\EndIf
\EndFor \\
\Return matches
\end{algorithmic}
\end{algorithm}

Algorithm \ref{alg-simple} shows the basic steps taken to find matches 
in two or more images. This algorithm is a simplified version of the 
proposed method completely agnostic of the geometry of the feature 
points in the picture. To better explain the algorithm here is a step by 
step explication of the inputs and functions mentioned:
\begin{itemize}
\item[]{\emph{images}: The images we are matching}
\item[]{\emph{$n$}: The number of weights per row in the adjecency 
matrix.}
\item[]{\emph{getFeatures}: Returns a set of feature points given an 
image. In practice the BRIEF\cite{calonder2010brief} feature point 
detector and descriptor has worked well, but it could easily be replaced 
with any other feature point type.} 
% TODO: include a comparison of SURF, SIFT, etc
\item[]{\emph{getAdjecencyMatrix}: Based on the distance measure used to 
compare the similarity of the feature points a symmetric adjecency 
matrix is returned.}
\item[]{\emph{pruneEdges}: The result of the Louvain clustering depends 
on the density of the graph. To ensure an optimal number of partitions 
we only keep the $n$ highest weights per row and column in the adjecency 
matrix, making sure it's still symmetric afterwards. In practice setting 
$n$ to around four or five yields good results.}
\item[]{\emph{cluster}: Cluster the graph corresponding to the adjecency 
matrix given using the Louvain algorithm in \cite{blondel2008}.}
\item[]{\emph{getPositions}: Given a trait consisting of features in 
several different images, find the positions of traits in each image and 
return them. This could either be done by taking a mean of the feature 
points in an image, or by returning the position of the highest matching 
feature points within the trait. In practice ... TODO}
\end{itemize}

\subsubsection{On the subject of features}

TODO: Show examples of the same image with feature points from different 
detectors and explain the difference. Go in to why brief seems to work 
best in practice and superficially explain how brief works compared with 
SIFT.

\begin{algorithm}
\caption{Cluster-Match Algorithm With Geometry}
\label{alg-geometry}
\begin{algorithmic}
\Require $images$ : set of images, $\alpha$ : float $\in \left[0, 
1\right]$, $\beta$ : float $\in \left[0, 1\right]$
\State $matches\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features of all images
	\State $f_i\gets getFeatures(I_i)$
\EndFor
\State $A\gets getAdjecencyMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $B\gets getGeomDistanceMatrix(f_1, f_2,\; \ldots \;, f_n)$
\State $M\gets getInterImageMask(f_1, f_2,\; \ldots \;, f_n)$
\State $B_{norm}\gets 1 - normalize(G)$ \State $A_{norm}\gets 1 - 
normalize(A)$
\State $A_{geom}\gets A_{norm}\left[ \neg M\right] + \beta \cdot 
B_{norm}\left[M\right]$
\State $A_{pruned}\gets pruneEdges(A_{geom},\alpha)$
\State $P\gets cluster(A_{pruned})$ 
\ForAll{$p \in P$}
	\State $trait\gets \left\{p \cap f_i \mid p \cap f_i \neq 
\varnothing, f_i \in \left\{f_1, f_2,\; \ldots \;, f_n\right\}\right\}$
	\If{$\left\vert trait \right\vert > 1$}
		\State $matches\gets getPositions(trait)$
		\Comment Position of trait in each image
	\EndIf
\EndFor \\
\Return matches
\end{algorithmic}
\end{algorithm}

While clustered feature points are often geometrically close together, 
there might be cases where several points in the same image are similar 
enough to all be included in one cluster despite their geometrical 
distance.  To solve this problem we can either modify the clustering 
algorithm to take the geometric distances into account like 
\cite{das2008event}.  Alternatively we can modify the weight matrix to 
penalize putting distanced feature points in the same cluster. However 
we only know the geometrical distance between two feature points as long 
as both belong to the same image. One way to approach this problem is to 
replace all the weights of edges between two feature points within the 
same image with their distance. For the case of two images the 
normalized adjecency matrix is defined as follows:
%
$$A_{norm}=
\begin{bmatrix}
w_{1,1}   & w_{1,2}   & \ldots & w_{1,n-1}   & w_{1,n}     \\
w_{2,1}   & w_{2,2}   & \ldots & w_{2,n-1}   & w_{2,n}     \\
\vdots    & \vdots    & \ddots & \vdots      & \vdots      \\
w_{n-1,1} & w_{n-1,2} & \ldots & w_{n-1,n-1} & w_{n-1,n}   \\
w_{n,1}   & w_{n,2}   & \ldots & w_{n,n-1}   & w_{n,n}
\end{bmatrix}=
\begin{bmatrix}
A_{im1} & S       \\
T       & A_{im2} \\
\end{bmatrix}
$$
%
In a similar manner we can define an adjecency matrix based on the 
geometrical distance. Note for example how $d_{1,n}$ can take any value 
as it denotes the geometric distance of feature points in two different 
images and as such can be set to any value.
%
$$B_{norm}=
\begin{bmatrix}
d_{1,1}   & d_{1,2}   & \ldots & d_{1,n-1}   & d_{1,n}     \\
d_{2,1}   & d_{2,2}   & \ldots & d_{2,n-1}   & d_{2,n}     \\
\vdots    & \vdots    & \ddots & \vdots      & \vdots      \\
d_{n-1,1} & d_{n-1,2} & \ldots & d_{n-1,n-1} & d_{n-1,n}   \\
d_{n,1}   & d_{n,2}   & \ldots & d_{n,n-1}   & d_{n,n}
\end{bmatrix}=
\begin{bmatrix}
B_{im1} & R       \\
Q       & B_{im2} \\
\end{bmatrix}
$$
%
Given these two and a weight factor $\beta$ we can construct a new 
weight matrix as follows:
%
$$C=
\begin{bmatrix}
\beta \cdot B_{im1} & S       \\
T       & \beta \cdot B_{im2} \\
\end{bmatrix}
$$
%
Since we prune the matrix afterwards, it's important to choose a weight 
factor that doesn't lead to all edges being either intra image edges or 
intra image edges, since the first would create no clusters across 
images and the latter would mean that the geometrical information would 
be discarded and the bipartite graph constituting of edges between 
feature points in the two images would be used instead.
